A tutorial of Bayesian beta regressions with brms in R

### 2024-12-13 manuscript/manuscript.qmd
add Paul to authors
edit manuscript

---
title: A tutorial of Bayesian beta regressions with brms in R
author:
  - name: Stefano Coretta
    email: s.coretta@ed.ac.uk
    affiliations: 
        - id: uoe
          name: University of Edinburgh
          department: Linguistics and English Language
          address: 3 George Sq
          city: Edinburgh
          state: United Kingdom
          postal-code: EH8 9AD
    attributes:
        corresponding: true
    note: This is the first author footnote.
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: linguistics.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
---

@@

```{r}
#| label: nas-prop-r
library(tidybayes)
nas_prop_r <- nas_prop_bm |> 
  spread_draws(r_speaker[speaker,voicingvowel]) |> 
  separate(voicingvowel, c("voicing", "vowel")) |> 
  left_join(y = nas_prop_bm_draws_long) |> 
  mutate(r_speaker_value = value + r_speaker) |> 
  group_by(.draw, speaker, voicing) |> 
  summarise(r_speaker_value_mean = mean(r_speaker_value)) |> 
  pivot_wider(names_from = voicing, values_from = r_speaker_value_mean) |> 
  mutate(voicing_diff = voicingvoiceless - voicingvoiced)
```

>>>>

---
title: A tutorial of Bayesian beta regressions with brms in R
author:
  - name: Stefano Coretta
    email: s.coretta@ed.ac.uk
    affiliations: 
        - id: uoe
          name: University of Edinburgh
          department: Linguistics and English Language
          address: 3 George Sq
          city: Edinburgh
          state: United Kingdom
          postal-code: EH8 9AD
    attributes:
        corresponding: true
    note: This is the first author footnote.
  - name: Paul Bürkner
    email: paul.buerkner@gmail.com
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: linguistics.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
---

@@

Moving onto question 2: is there individual speaker variation? [TAMMINGA XXX] For this, we will use the `spread_draws()` function from tidybayes [XXX] to extract the draws of the varying terms (in brms these are the coefficients that start with `r_`). There is quite a few steps of processing to get from the raw draws to the estimand we need: while we have commented the following code, we encourage readers to test each line sequentialy and inspect the intermediate output to fully understand the process.

```{r}
#| label: nas-prop-r
library(tidybayes)

nas_prop_r <- nas_prop_bm |> 
  # extract draws from model, only `r_speaker` varying terms
  spread_draws(r_speaker[speaker,voicingvowel]) |> 
  # separate the column voicingvowel to two columns
  separate(voicingvowel, c("voicing", "vowel")) |> 
  # join the draws with the `b_` terms
  left_join(y = nas_prop_bm_draws_long) |> 
  # get the expected log-odd value of each speaker, in each draw
  # this is the sum of the `value` from the b_ terms and the value from the
  # r_speaker term.
  mutate(r_speaker_value = value + r_speaker) |> 
  # group the data for summarise
  group_by(.draw, speaker, voicing) |> 
  # get mean expected log-odds by draw, speaker and voicing (averaging across vowel)
  summarise(r_speaker_value_mean = mean(r_speaker_value)) |> 
  # make the data wider: two columns, one for voiced and one for voiceless
  pivot_wider(names_from = voicing, values_from = r_speaker_value_mean) |> 
  # finally, calculate the difference in expected log-odds of voiceless and voiced
  mutate(voicing_diff = voicingvoiceless - voicingvoiced)

nas_prop_r
```

@fig-nas-prop-r plots the posterior distributions of the expected log-odd difference of coarticulatory nasalisation in voiceless vs voiced NC sequences (*x*-axis), for each speaker in the data (*y*-axis), as predicted by the model. The red solid vertical line indicates the constant (overall) expected log-odd difference based on the (constant) `b_` terms. The black dashed vertical line marks log-difference 0 (i.e., no difference in proportion of nasalisation between voiceless and voiced NC). There is a lot of uncertainty within and between speakers: while the distributions of most speakers are located in the positive range, some expected distributions (see bottom of figure) do span both negative and positive values. In other words, while most speakers are expected to have a larger nasalisation proportion in voiceless NC sequences, a few in fact might have the opposite pattern. Even among those speakers that do have a positive difference, there is a lot of uncertainty as to the magnitude of the difference.
